Mechanisms of detachment in fibrillar adhesive systems

### 2021-05-06 review/manuscript_HJB.tex

Over the past few decades, there has been numerous studies to understand
how animals such as geckos and insects are able to walk on surfaces
of any direction while defying gravity. A microscopic observation
reveals that, in many cases, animals have a dense array of hairy structures
at the end of their legs \cite{RN198,RN59}. These hairy adhesive
pads help the animal to stay attached to any surface or detach easily
at will for countless cycles, a property that is referred to as \emph{reversible
adhesion}. Previous attempts to theoretically explain adhesion in
hairy pads\cite{RN40,RN120} has followed two fundamental approaches:
either by energy balance, or by force balance.

In the energy balance approach, adhesion is characterized by ``work
of adhesion'' (W\textsubscript{adh}) which is the energy required
to separate a pad from the surface. During detachment, the elastic
energy stored in the hair is dissipated, which increases W\textsubscript{adh}
and thus adhesion is enhanced\cite{RN29,RN114}. Detachment of an
individual hair can be explained based on Kendall's peeling theory\cite{RN147},
which predicts low forces at high peeling angles.

In the force balance approach, adhesion is characterized by pull-off
force, F\textsubscript{p} (or stress, \textgreek{sv}\textsubscript{p}),
which is the minimum force necessary to separate two surfaces from
contact. Based on a ``cohesive zone model'', Hui et. al.\cite{RN72}
had identified two regimes of single hair detachment: 1) a flaw sensitive
regime, where contact failure occurs due to crack propagation leading
to low \textgreek{sv}\textsubscript{p}, 2) a flaw insensitive regime,
where the contact interface fails simultaneously, leading to high
\textgreek{sv}\textsubscript{p}. Tian et. al. \cite{RN230} had shown
that the ``spatula'' shaped hair tips in a gecko's toe allows it
to change adhesion by three orders of magnitude by laterally sliding
and controlling the pulling angle to disorient the hairs. Federle\cite{RN20}
had further argued that the curved shape of the hair helps the pad
to stay attached when pulled proximally, and easily detached by elastic
recoil when pushed distally.

The theories presented so far suggests that low detachment force of
a hairy pad can be attained by 1) increasing the stress concentration
by peeling the pad at high angles, or 2) laterally shearing the pad
before pull-off, which requires the hairs to have an asymmetric geometry
or curvature. However, some insects like male dock beetles predominantly
possess flat ``mushroom'' tipped hairs on their pads\cite{RN19},
which are relatively less asymmetric compared to the ``spatula''
tipped hairs. This begs the question of how the beetle can still easily
detach their legs during locomotion, since these mushroom shaped hairs
should be resilient to detachment via lateral shearing. Artificial
micro-pillar adhesives mimicking the mushroom shaped hairs of the
beetle have in fact demonstrated its superior adhesive strength in
comparison to other geometries\cite{RN112,RN57}. Introducing asymmetry
into the pillar geometry of such mimics for easy detachment is however
challenging due to current limitations in fabrication techniques and
difficulty in scaling-up for practical applications \cite{RN57}.
Alternate strategies are thus desired to switch the adhesion state
of symmetric pillar arrays in a reversible manner, e.g. by buckling
the pillars under compressive load leading to their contact loss \cite{RN279}
or by using special materials reacting to external stimuli such as
magnetic field\cite{RN186}, UV light \cite{RN280} or temperature\cite{RN281}.

In this work, we propose the possible mechanisms by which adhesive
pads with symmetric hairs can be easily detached under a theoretical
framework. The hairy pad is modelled as an array of springs and the
force-balance approach is used to quantify the net adhesion \cite{RN117}.
We hope our work to provide strategies to control the adhesion force
of an artificial micro-pillar adhesive, which has applications in
bio-inspired climbing robots, pick-and-place operations and reusable
adhesives.

@@

The hairy pad is assumed to be a one dimensional array of $N_{t}$
hairs, each behaving like a spring with spring constant, $k_{h}$,
and natural length, $l_{h,0}$ (Figure \ref{fig:Schematic}). The
array backing is assumed to be stiff. The pad is attached to a linearly
deformable leg, assumed to be another spring with spring constant,
$k_{l}$, and natural length, $l_{l,0}$. The leg is hinged to the
array at a distance, $s$, from the right end of the array. The hinge
is at a vertical distance, $d_{s}$, from the surface. The hairs are
spaced apart by a width, $w$, and the array is of length, $L=\left(N_{t}-1\right)w$.
The pad is oriented at an angle, $\theta$, while making contact with
a flat smooth surface. Each hair can attain a maximum length, $l_{h,p}$,
before pull-off, such that its pull-off force, $f_{p}=k_{h}\left(l_{h,p}-l_{h,0}\right)$.
$F_{net}$ is the net force on the pad and $M_{net}$ is the the net
torque about the hinge, at a particular instant during the detachment
process.

@@

Equation \ref{eq:force_n_d} will be valid for $d_{s}\leq d_{s,max}$.

The maximum possible adhesion of the array would be the case when
all hairs detach simultaneously ($\theta=0\lyxmathsym{\textdegree}$):

@@

The net torque, $M_{net}$, about the hinge can be similarly derived
(see Appendix):

>>>>

Over the past few decades, there has been numerous studies to understand
how animals such as geckos and insects are able to walk on surfaces
of any direction while defying gravity. A microscopic observation
reveals that, in many cases, animals have a dense array of hairy structures
at the end of their legs \cite{RN198,RN59}. These hairy adhesive
pads help the animal to stay attached to 
surfaces or detach easily
at will for countless cycles, a property that is referred to as \emph{reversible
adhesion}. Previous attempts to theoretically explain adhesion in
hairy pads\cite{RN40,RN120} has followed two fundamental approaches:
either by energy balance, or by force balance.

In the energy balance approach, adhesion is characterized by ``work
of adhesion'' (W\textsubscript{adh}) which is the energy required
to separate a pad from the surface. During detachment, the elastic
energy stored in the hair is dissipated, which increases W\textsubscript{adh}
and thus adhesion is enhanced\cite{RN29,RN114}. 
% I do not understand the last sentence. What precisely does "`adhesion is enhanced"' mean? If the work of adhesion is fixed, how can it be adhanced? 
Detachment of an
individual hair can be explained based on Kendall's peeling theory\cite{RN147},
which predicts low forces at high peeling angles.
% The last sentence seems to belonmg more to the next paragraph.

In the force balance approach, adhesion is characterized by a pull-off
force, F\textsubscript{p} (or stress, \textgreek{sv}\textsubscript{p}),
which is the minimum force necessary to separate two surfaces from
contact. Based on a ``cohesive zone model'', Hui et. al.\cite{RN72}
had identified two regimes of single hair detachment: 1) a flaw sensitive
regime, where contact failure occurs due to crack propagation leading
to low \textgreek{sv}\textsubscript{p}, 2) a flaw insensitive regime,
where the contact interface fails simultaneously, leading to high
\textgreek{sv}\textsubscript{p}. Tian et. al. \cite{RN230} had shown
that the ``spatula'' shaped hair tips in a gecko's toe allows it
to change adhesion by three orders of magnitude by laterally sliding
and controlling the pulling angle to disorient the hairs. Federle\cite{RN20}
had further argued that the curved shape of the hair helps the pad
to stay attached when pulled proximally, and easily detached by elastic
recoil when pushed distally.

% I suggest to describe both models a bit more in detail. What is a flaw sensitive regime? What does "`flaw"' mean and how is it related to adhesion?

The theories presented so far suggests that low detachment force of
a hairy pad can be attained by 1) increasing the stress concentration
by peeling the pad at high angles, or 2) laterally shearing the pad
before pull-off, which requires the hairs to have an asymmetric geometry
or curvature. However, some insects like male dock beetles predominantly
possess flat ``mushroom'' tipped hairs on their pads\cite{RN19}.
They are relatively symmetric compared to the ``spatula''
tipped hairs. This begs the question of how the beetle can still easily
detach their legs during locomotion, since these mushroom shaped hairs
should be resilient to detachment via lateral shearing. Artificial
micro-pillar adhesives mimicking the mushroom shaped hairs of the
beetle have in fact demonstrated its superior adhesive strength in
comparison to other geometries\cite{RN112,RN57}. Introducing asymmetry
into the pillar geometry of such mimics for easy detachment is however
challenging due to current limitations in fabrication techniques and
difficulty in scaling-up for practical applications \cite{RN57}.
Alternate strategies are thus desired to switch the adhesion state
of symmetric pillar arrays in a reversible manner, e.g. by buckling
the pillars under compressive load leading to their contact loss \cite{RN279}
or by using special materials reacting to external stimuli such as
magnetic field\cite{RN186}, UV light \cite{RN280} or temperature\cite{RN281}.

In this work, we propose a mechanisms by which adhesive
pads with symmetric hairs can easily detached from a surface. 
The hairy pad is modeled as an array of springs and the
force-balance approach is used to quantify the net adhesion \cite{RN117}.
We hope our work to provide strategies to control the adhesion force
of an artificial micro-pillar adhesive, which has applications in
bio-inspired climbing robots, pick-and-place operations and reusable
adhesives.

@@

The hairy pad is assumed to be a one dimensional array of $N_{t}$
hairs, each behaving like a spring with spring constant, $k_{h}$,
and natural length, $l_{h,0}$ (Figure \ref{fig:Schematic}). The
array backing is assumed to be stiff. The pad is attached to a linearly
deformable leg, assumed to be another spring with spring constant,
$k_{l}$, and natural length, $l_{l,0}$. The leg is hinged to the
array at a distance, $s$, from the right end of the array. The hinge
is at a vertical distance, $d_{s}$, from the surface. The hairs are
spaced apart by a width, $w$, and the array is of length, $L=\left(N_{t}-1\right)w$.
The pad is oriented at an angle, $\theta$, while making contact with
a flat smooth surface. Each hair can attain a maximum length, $l_{h,p}$,
before pull-off, such that its pull-off force, $f_{p}=k_{h}\left(l_{h,p}-l_{h,0}\right)$.
$F_{net}$ is the net force on the pad and $M_{net}$ is the the net
torque about the hinge, at a particular instant during the detachment
process.
% Is the angle at the hinge fixed? What happens, when I start pulling vertically?

@@

Equation \ref{eq:force_n_d} will be valid for $d_{s}\leq d_{s,max}$. 
The maximum possible adhesion of the array would be the case when
all hairs detach simultaneously ($\theta=0\lyxmathsym{\textdegree}$):

@@

The net torque, $M_{net}$, about the hinge can be similarly derived
(Appendix):
